---
title: "Logistic Regression analysis of the credit data"
Question: 1
output: html_notebook
---
#Step 1 - Collecting data
```{r}
credit <- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml7/credit.csv")

```

#Step 2 - Exploring and preparing the data
```{r}
## Fix the default variable to be 0 or 1

credit$default <- as.numeric(credit$default)
credit$default <- credit$default - 1

# examine the launch data
str(credit)

```

```{r}
# logisitic regression

# set up trainning and test data sets

indx <- sample(1:nrow(credit), as.integer(0.9*nrow(credit)))
indx

credit_train <- credit[indx,]
credit_test <- credit[-indx,]

```


```{r}
# Check if there are any missing values

library(Amelia)
missmap(credit, main = "Missing values vs observed")

# number of missing values in each column

sapply(credit,function(x) sum(is.na(x)))

# number of unique values in each column

sapply(credit, function(x) length(unique(x)))
```
#There are no missing values in the dataset

#Step 3 - Training the Logistic regression model on the data

```{r}
# fit the logistic regression model, with all predictor variables

model <- glm(default ~.,family=binomial(link='logit'),data=credit_train)
#argument family = binomial tells R to run a logistic regression
model

summary(model)

anova(model, test="Chisq")

```


```{r}
# drop the insignificant predictors, alpha = 0.10
#Clearly, checking_balance, months_loan_duration, credit_history,  percent_of_income and age are significant predictors (marked with *)

model <- glm(default ~ checking_balance + months_loan_duration + credit_history +  percent_of_income + age,family=binomial(link='logit'),data=credit_train)
model
```

#Step 4 - Evaluating model performance

```{r}
summary(model)

anova(model, test="Chisq")
```


```{r}
# check Accuracy

fitted.results <- predict(model,newdata=credit_test,type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)

misClasificError <- mean(fitted.results != credit_test$default)
print(paste('Accuracy',1-misClasificError))
```
#The 0.79 accuracy on the test set is pretty good result

#Step 5 - Improving model performance

```{r}
# ROC

library(ROCR)
p <- predict(model, newdata=credit_test, type="response")
pr <- prediction(p, credit_test$default)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)
```

```{r}
#auc

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```
# Area under the curve (auc) is 0.81 i.e. closer to 1. We can say that model has a good predictive ability.


#Random Forest analysis of credit data***************
#Question: 2


#Step 1 - Collecting data
```{r}
credit <- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml7/credit.csv")

barplot(table(credit$default))
```

#There are a lot of declined credit  as compared to accepted. 

#Step 2 - Exploring and preparing the data
```{r}
str(credit)

```


```{r}
# Random Forest regression

# Before we build our model, let’s separate our data into testing and training sets

indx <- sample(1:nrow(credit), (0.7*nrow(credit)))
indx

credit_train <- credit[indx,]
credit_test <- credit[-indx,]

```
```{r}
# Check if there are any missing values

library(Amelia)
missmap(credit, main = "Missing values vs observed")

# number of missing values in each column

sapply(credit,function(x) sum(is.na(x)))

# number of unique values in each column

sapply(credit, function(x) length(unique(x)))
```
#No missing values in the dataset

#Step 3 - Training the Random Forest model on data

```{r}
### Building the model

library(randomForest)

model <- randomForest(default ~., data = credit_train, ntree=1000, mtry=8)
```

```{r}
model
```

```{r}
pred <- predict(model, newdata = credit_test)
```

#Step 4 - Evaluating model performance
#Confusion Matrix

```{r}
table(pred, credit_test$default)
```

#Accuracy

```{r}
sum(pred==credit_test$default) / nrow(credit_test)
```

#We achieved 77% accuracy with random forest model. It could be further improved by feature selection, and possibly by trying different values of mtry.
